Supporting information: Copulas and their potential for ecology

### 2019-10-01 Paper.Rmd
table and figure simplified as suggested by DCR

Our model with asymmetric environmental effects 
produces outputs with visually apparent asymmetry 
of tail dependence, to an extent that depended on the value of $b$;
thus this is a mechanism that can produce non-normal copula structure and
asymmetric tail dependence across space in ecological variables.
<!--***DAN: move the following to the end of this part-->
This result is represented in Fig. \ref{MasterFigure} as the solid box around 
"Nonlinear environmental effects, Liebig's law" and the arrows labelled "A".
It is explained in the Discussion why the box and some of those arrows 
are solid, instead of dashed, although the results we 
present in this section are theoretical.
<!--***DAN: move the above to the end of this part-->
For $b=0.1$ (Fig. \ref{fig_ms_asym_b_0.1}) and $b=-0.1$ 
(Fig. \ref{SM-fig_ms_asym_b_-0.1}), 
for both $g_1$ and $g_2$,
asymmetry of tail dependence was strong; for 
$b=\pm 0.5$, asymmetry was weaker but still apparent
(Figs \ref{SM-fig_ms_asym_b_0.5} - 
\ref{SM-fig_ms_asym_b_-0.5}).
<!--***DAN: Shyamolina, can you please combine Figs S24-S26 into one single figure
with 6 panels? Please also search for citations in the main text and sup mat that 
used to refer to the three figures and replace them with a citation to the
new, combined figure.-->
Lower-tail (respectively, upper-tail) spatial dependence in 
effects of noise ($g_1$, respectively, $g_2$) 
produced lower-tail (respectively, upper-tail) spatial 
dependence in the ecological variable.
Model selection results (Table \ref{SM-tab_summary_stat_asym_sens}) confirmed that
ecological outputs had non-normal copula structure, but did not always reflect 
asymmetry of tail dependence, perhaps because those methods measured
dependence in the extreme tails. Results using our nonparametric statistics, however,
reflected the visually apparent asymmetry (Tables \ref{SM-tab_stat_npa_asym}). 
<!--***DAN: Simplify as follows, maybe:
1) get rid of table S10 (you will have to change the methods so it no longer says we
do our model selection analysis here)
2) simplify table S11 so it only shows the cor results (also means a change to methods)
and then maybe bring the resulting table to main text?
-->

<!--***DAN: Shyamolina, can you please simplify Table S11 so it only shows the cor
results? That also means the third column will be unnecessary. Please do this as follows:
1) keep but comment out the code that produces the old table; 2) write new code (presumably
starting from a copy of the old code) that makes the new, simpler version of the table. 
That will make it very easy to produce the old version again if we ever want to.-->

>>>>

Our model with asymmetric environmental effects 
produces outputs with visually apparent asymmetry 
of tail dependence, to an extent that depended on the value of $b$;
thus this is a mechanism that can produce non-normal copula structure and
asymmetric tail dependence across space in ecological variables.
<!--***DAN: move the following to the end of this part-->
This result is represented in Fig. \ref{MasterFigure} as the solid box around 
"Nonlinear environmental effects, Liebig's law" and the arrows labelled "A".
It is explained in the Discussion why the box and some of those arrows 
are solid, instead of dashed, although the results we 
present in this section are theoretical.
<!--***DAN: move the above to the end of this part-->
For $b=0.1$ (Fig. \ref{fig_ms_asym_b_0.1}) and $b=-0.1$ 
(Fig. \ref{SM-fig_ms_asym_b_-0.1_0.5_-0.5}A,B), 
for both $g_1$ and $g_2$,
asymmetry of tail dependence was strong; for 
$b=\pm 0.5$, asymmetry was weaker but still apparent
(Fig. \ref{SM-fig_ms_asym_b_-0.1_0.5_-0.5}C-F).
<!--***DAN: Shyamolina, can you please combine Figs S24-S26 into one single figure
with 6 panels? Please also search for citations in the main text and sup mat that 
used to refer to the three figures and replace them with a citation to the
new, combined figure.
***Shya: Done.-->
Lower-tail (respectively, upper-tail) spatial dependence in 
effects of noise ($g_1$, respectively, $g_2$) 
produced lower-tail (respectively, upper-tail) spatial 
dependence in the ecological variable.
Model selection results (Table \ref{SM-tab_summary_stat_asym_sens}) confirmed that
ecological outputs had non-normal copula structure, but did not always reflect 
asymmetry of tail dependence, perhaps because those methods measured
dependence in the extreme tails. Results using our nonparametric statistics, however,
reflected the visually apparent asymmetry (Tables \ref{SM-tab_stat_npa_asym}). 
<!--***DAN: Simplify as follows, maybe:
1) get rid of table S10 (you will have to change the methods so it no longer says we
do our model selection analysis here)
2) simplify table S11 so it only shows the cor results (also means a change to methods)
and then maybe bring the resulting table to main text?
-->

<!--***DAN: Shyamolina, can you please simplify Table S11 so it only shows the cor
results? That also means the third column will be unnecessary. Please do this as follows:
1) keep but comment out the code that produces the old table; 2) write new code (presumably
starting from a copy of the old code) that makes the new, simpler version of the table. 
That will make it very easy to produce the old version again if we ever want to.
***Shya: Done.-->

### 2019-10-01 SupportingInformation.Rmd
table and figure simplified as suggested by DCR

pdf("./Results/asym_sens_results/a_0.2_b_0.5_r_0.8_asym_sens.pdf",height=3,width=6)
op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.5,0))
pop<-ans_b_0.5$pop
plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
     cex.axis=1,col=rgb(0,0,0,0.2),pch=20,cex=0.6)
text(0.1,0.95,"A",cex=2)
pop<-ans_b_0.5$pop_d
plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
     cex.axis=1,col=rgb(0,0,0,0.2),pch=20,cex=0.6)
text(0.1,0.95,"B",cex=2)
par(op)
dev.off()

pdf("./Results/asym_sens_results/a_0.2_b_-0.5_r_0.8_asym_sens.pdf",height=3,width=6)
op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.5,0))
pop<-ans_b_neg0.5$pop
plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
     cex.axis=1,col=rgb(0,0,0,0.2),pch=20,cex=0.6)
text(0.1,0.95,"A",cex=2)
pop<-ans_b_neg0.5$pop_d
plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
     cex.axis=1,col=rgb(0,0,0,0.2),pch=20,cex=0.6)
text(0.1,0.95,"B",cex=2)
par(op)
dev.off()
```

<!--adding plots from asymmetric dependence analysis results-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_-0.1_r_0.8_asym_sens.pdf}
\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=-0.1$.\label{fig_ms_asym_b_-0.1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_0.5_r_0.8_asym_sens.pdf}
\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=0.5$.\label{fig_ms_asym_b_0.5}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_-0.5_r_0.8_asym_sens.pdf}
\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=-0.5$.\label{fig_ms_asym_b_-0.5}}
\end{center}
\end{figure}

<!--Displaying summary table with model selection results from asymmetric dependence analysis-->
```{r tab_summary_stat_asym_sens,echo=F,results="markup"}
summary_stat<-summary_stat[c(1:6),]
rownames(summary_stat)<-c("Best-fit copula AICw","Normal copula AICw","Model-avg. LT","Model-avg. UT","p, CvM","p, KS")
knitr::kable(summary_stat, booktabs = T, escape=F,
             format = "latex", linesep = "",
             digits=4, col.names=c("b=0.1, g$_{1}$","b=0.1, g$_{2}$",
                                   "b=-0.1, g$_{1}$","b=-0.1, g$_{2}$",
                                   "b=0.5, g$_{1}$","b=0.5, g$_{2}$",
                                   "b=-0.5, g$_{1}$","b=-0.5, g$_{2}$"),
      caption.short="Asymmetric sensitivity model, model selection results",
      caption="Asymmetric sensitivity model, model selection results. 
The output of simulations from the first model detailed in 
the methods section for Q2 (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
\\ref{fig_ms_asym_b_-0.1}-\\ref{fig_ms_asym_b_-0.5}) were subjected to the model 
selection procedures detailed in the 
methods section for Q1. Best-fit copula AICw: the AIC weight of the best-fitting copula family;
Normal copula AICw: the AIC weight of the normal copula family; Model-avg. LT: the lower-tail 
statistic, weighted-averaged across models using AICw; Model-avg. UT: same, but for the 
upper-tail statistic; p, CvM: p-value for goodness of fit test of the best-fitting copula
family using the test based on the Cramer-von Mises statistic; p, KS: p-value for goodness of fit 
test of the best-fitting copula family using the test based on the Kolmogorov-Smirnov 
statistic.\\label{tab_summary_stat_asym_sens}")
```

<!--Displaying summary table with non-parametric approach results of asymmetric dependence analysis-->
```{r tab_stat_npa_asym,echo=F,results="markup"}
library(kableExtra)

tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.5.RDS")
tab_stat_npa_asym<-c()
for(i in c(1:8)){
  temp<-make_table_stat_fn(data=tab_asym[[i]])
  tab_stat_npa_asym<-rbind(tab_stat_npa_asym,temp)
}
gtext<-c("b=0.1, g$_{1}$","b=0.1, g$_{2}$","b=-0.1, g$_{1}$","b=-0.1, g$_{2}$","b=0.5, g$_{1}$","b=0.5, g$_{2}$","b=-0.5, g$_{1}$","b=-0.5, g$_{2}$")
tab_stat_npa_asym<-cbind(b=rep(c(0.1,0.1,-0.1,-0.1,0.5,0.5,-0.5,-0.5),each=3),
  g=rep(c("g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$"),each=3),tab_stat_npa_asym)
knitr::kable(tab_stat_npa_asym, booktabs = T, longtable=T,
             format = "latex",linesep = "",escape=F,
             caption.short = "Asymmetric sensitivity model, nonparametric statistics results",
             caption ="Asymmetric sensitivity model, nonparametric statistics results. 
The output of simulations from the first model detailed in 
the methods section for Q2 (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
\\ref{fig_ms_asym_b_-0.1}-\\ref{fig_ms_asym_b_-0.5}) were subjected to nonparamtric statistical 
analyses decribed in the methods section for Q1. The listed statistic was computed for the
output of the model with the indicated b and g, and the value was compared to 1000 values
of the same statistic computed on surrogate time series randomized to have normal copula structure
(see methods for Q1). A table entry $<X$ (respectively, $>X$) indicates the value of the 
given statistic on the data was less (respectively, more) than its value on $X$ of the 
surrogates, so entries of the form $<X$ (respectively, $>X$) for $X$ equal to $975$ or above 
indicate that upper-tail (respectively, lower-tail) dependence was significantly stronger than 
lower-tail dependence compared to a normal-copula null hypothesis. `FS' stands for `first slice'
and refers to the bounds 0 to 0.1; `LS' stands for `last slice' and refers to the bounds
0.9 to 1.\\label{tab_stat_npa_asym}")
```

>>>>

pdf("./Results/asym_sens_results/a_0.2_b_0.5_r_0.8_asym_sens.pdf",height=3,width=6)
op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.5,0))
pop<-ans_b_0.5$pop
plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
     cex.axis=1,col=rgb(0,0,0,0.2),pch=20,cex=0.6)
text(0.1,0.95,"C",cex=2)
pop<-ans_b_0.5$pop_d
plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
     cex.axis=1,col=rgb(0,0,0,0.2),pch=20,cex=0.6)
text(0.1,0.95,"D",cex=2)
par(op)
dev.off()

pdf("./Results/asym_sens_results/a_0.2_b_-0.5_r_0.8_asym_sens.pdf",height=3,width=6)
op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.5,0))
pop<-ans_b_neg0.5$pop
plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
     cex.axis=1,col=rgb(0,0,0,0.2),pch=20,cex=0.6)
text(0.1,0.95,"E",cex=2)
pop<-ans_b_neg0.5$pop_d
plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
     cex.axis=1,col=rgb(0,0,0,0.2),pch=20,cex=0.6)
text(0.1,0.95,"F",cex=2)
par(op)
dev.off()
```

<!--adding plots from asymmetric dependence analysis results-->

\begin{figure}[!h]
\begin{center}
\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_-0.1_r_0.8_asym_sens.pdf}
\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_0.5_r_0.8_asym_sens.pdf}
\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_-0.5_r_0.8_asym_sens.pdf}
\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=-0.1$ (A-B), $b=0.5$ (C-D) and $b=-0.5$ (E-F).\label{fig_ms_asym_b_-0.1_0.5_-0.5}}
\end{center}
\end{figure}

<!--Displaying summary table with model selection results from asymmetric dependence analysis-->
```{r tab_summary_stat_asym_sens,echo=F,results="markup"}
summary_stat<-summary_stat[c(1:6),]
rownames(summary_stat)<-c("Best-fit copula AICw","Normal copula AICw","Model-avg. LT","Model-avg. UT","p, CvM","p, KS")
knitr::kable(summary_stat, booktabs = T, escape=F,
             format = "latex", linesep = "",
             digits=4, col.names=c("b=0.1, g$_{1}$","b=0.1, g$_{2}$",
                                   "b=-0.1, g$_{1}$","b=-0.1, g$_{2}$",
                                   "b=0.5, g$_{1}$","b=0.5, g$_{2}$",
                                   "b=-0.5, g$_{1}$","b=-0.5, g$_{2}$"),
      caption.short="Asymmetric sensitivity model, model selection results",
      caption="Asymmetric sensitivity model, model selection results. 
The output of simulations from the first model detailed in 
the methods section for Q2 (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
\\ref{fig_ms_asym_b_-0.1_0.5_-0.5}) were subjected to the model 
selection procedures detailed in the 
methods section for Q1. Best-fit copula AICw: the AIC weight of the best-fitting copula family;
Normal copula AICw: the AIC weight of the normal copula family; Model-avg. LT: the lower-tail 
statistic, weighted-averaged across models using AICw; Model-avg. UT: same, but for the 
upper-tail statistic; p, CvM: p-value for goodness of fit test of the best-fitting copula
family using the test based on the Cramer-von Mises statistic; p, KS: p-value for goodness of fit 
test of the best-fitting copula family using the test based on the Kolmogorov-Smirnov 
statistic.\\label{tab_summary_stat_asym_sens}")
```

<!--*** Keeping old complex table for future use:
Displaying summary table with non-parametric approach results of asymmetric dependence analysis
```{r tab_stat_npa_asym,echo=F,results="markup"}
library(kableExtra)

tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.5.RDS")
tab_stat_npa_asym<-c()
for(i in c(1:8)){
  temp<-make_table_stat_fn(data=tab_asym[[i]])
  tab_stat_npa_asym<-rbind(tab_stat_npa_asym,temp)
}
gtext<-c("b=0.1, g$_{1}$","b=0.1, g$_{2}$","b=-0.1, g$_{1}$","b=-0.1, g$_{2}$","b=0.5, g$_{1}$","b=0.5, g$_{2}$","b=-0.5, g$_{1}$","b=-0.5, g$_{2}$")
tab_stat_npa_asym<-cbind(b=rep(c(0.1,0.1,-0.1,-0.1,0.5,0.5,-0.5,-0.5),each=3),
  g=rep(c("g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$"),each=3),tab_stat_npa_asym)
knitr::kable(tab_stat_npa_asym, booktabs = T, longtable=T,
             format = "latex",linesep = "",escape=F,
             caption.short = "Asymmetric sensitivity model, nonparametric statistics results",
             caption ="Asymmetric sensitivity model, nonparametric statistics results. 
The output of simulations from the first model detailed in 
the methods section for Q2 (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
\\ref{fig_ms_asym_b_-0.1_0.5_-0.5}) were subjected to nonparamtric statistical 
analyses decribed in the methods section for Q1. The listed statistic was computed for the
output of the model with the indicated b and g, and the value was compared to 1000 values
of the same statistic computed on surrogate time series randomized to have normal copula structure
(see methods for Q1). A table entry $<X$ (respectively, $>X$) indicates the value of the 
given statistic on the data was less (respectively, more) than its value on $X$ of the 
surrogates, so entries of the form $<X$ (respectively, $>X$) for $X$ equal to $975$ or above 
indicate that upper-tail (respectively, lower-tail) dependence was significantly stronger than 
lower-tail dependence compared to a normal-copula null hypothesis. `FS' stands for `first slice'
and refers to the bounds 0 to 0.1; `LS' stands for `last slice' and refers to the bounds
0.9 to 1.\\label{tab_stat_npa_asym}")
```
-->

<!---Simpler version: Displaying summary table with non-parametric approach results of asymmetric dependence analysis--->
```{r tab_stat_npa_asym_simple,echo=F,results="markup"}
library(kableExtra)

tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.5.RDS")
tab_stat_npa_asym<-c()
for(i in c(1:8)){
  temp<-make_table_stat_fn(data=tab_asym[[i]])
  tab_stat_npa_asym<-rbind(tab_stat_npa_asym,temp)
}

ind_corstat<-which(tab_stat_npa_asym$Statistic %in% c("cor$_{FS}$-cor$_{LS}$"))
tab_corstat_asym<-tab_stat_npa_asym[ind_corstat,]
tab_corstat_asym<-tab_corstat_asym[,-1]

#gtext<-c("b=0.1, g$_{1}$","b=0.1, g$_{2}$","b=-0.1, g$_{1}$","b=-0.1, g$_{2}$","b=0.5, g$_{1}$","b=0.5, g$_{2}$","b=-0.5, #g$_{1}$","b=-0.5, g$_{2}$")
#tab_stat_npa_asym<-cbind(gtext,tab_corstat_asym)

tab_corstat_asym<-cbind(b=c(0.1,0.1,-0.1,-0.1,0.5,0.5,-0.5,-0.5),
  g=c("g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$"),tab_corstat_asym)

rownames(tab_corstat_asym)<-NULL

knitr::kable(tab_corstat_asym, booktabs = T, 
             format = "latex",linesep = "",escape=F, 
             caption.short = "Asymmetric sensitivity model, nonparametric statistics results",
             caption ="Asymmetric sensitivity model, nonparametric statistics results. 
The output of simulations from the first model detailed in 
the methods section for Q2 (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
\\ref{fig_ms_asym_b_-0.1_0.5_-0.5}) were subjected to nonparamtric statistical 
analyses decribed in the methods section for Q1. The statistic ($cor_{FS} - cor_{LS}$) was computed for the
output of the model with the indicated b and g, and the value was compared to 1000 values
of the same statistic computed on surrogate time series randomized to have normal copula structure
(see methods for Q1). A table entry $<X$ (respectively, $>X$) indicates the value of the 
given statistic on the data was less (respectively, more) than its value on $X$ of the 
surrogates, so entries of the form $<X$ (respectively, $>X$) for $X$ equal to $975$ or above 
indicate that upper-tail (respectively, lower-tail) dependence was significantly stronger than 
lower-tail dependence compared to a normal-copula null hypothesis. `FS' stands for `first slice'
and refers to the bounds 0 to 0.1; `LS' stands for `last slice' and refers to the bounds
0.9 to 1.\\label{tab_stat_npa_asym}")
```
